---
title: 307_visualize_fig4-5.qmd
subtitle: Clouds modulate light-pollution effects on avian behavior
toc: true
format:
  html:
    embed-resources: false
    cache: false
abstract: This script visualizes Figures 4, 5, and Supp Figs. 1 and 2
date: 2026-09-12 # last modified
date-format: "yyyy-MM-dd"
---


# setup env

```{r}
#| label: setup-env
pacman::p_load(
  data.table,
  tidyverse,
  here,
  glmmTMB,
  patchwork,
  MetBrewer,
  ggh4x
)

here::i_am("scripts/L3/307_visualize_select_fig4.qmd")
```

# read data


```{r}
#| label: read-data
dir_304_final_data <- here("data", "L3", "304_final_data")
dir_305_models <- here("data", "L3", "305_models")
dir_307_fig4 <- here("data", "L3", "307_fig4")

## Occurrence data:
diurn_on <- readRDS(here(dir_304_final_data, "304d_diurn_on_final_50.rds"))
diurn_ev <- readRDS(here(dir_304_final_data, "304e_diurn_ev_final_50.rds"))
noc <- readRDS(here(dir_304_final_data, "304f_noc_final_50.rds"))


## log transform and scale:
diurn_on$alan_sc <- scale(log1p(diurn_on$avg_rad))
diurn_on$cld_sc <- scale(diurn_on$sunrise_cloud_cover_percent)
diurn_on$family <- as.factor(diurn_on$family)
diurn_on$precip_sc <- scale(diurn_on$sunrise_precipitation_mm)

diurn_ev$alan_sc <- scale(log1p(diurn_ev$avg_rad))
diurn_ev$cld_sc <- scale(diurn_ev$sunset_cloud_cover_percent)
diurn_ev$family <- as.factor(diurn_ev$family)
diurn_ev$precip_sc <- scale(diurn_ev$sunset_precipitation_mm)

noc$alan_sc <- scale(log1p(noc$avg_rad))
noc$cloud_sc <- scale(noc$nightly_cloud_cover_percent)
noc$family <- as.factor(noc$family)
noc$precip_sc <- scale(noc$nightly_precipitation_mm)

## Calculate scaled 0 mm precip:
precip_0_on <- (0 - attr(diurn_on$precip_sc, "scaled:center")) /
  attr(diurn_on$precip_sc, "scaled:scale")
precip_0_ev <- (0 - attr(diurn_ev$precip_sc, "scaled:center")) /
  attr(diurn_ev$precip_sc, "scaled:scale")
precip_0_noc <- (0 - attr(noc$precip_sc, "scaled:center")) /
  attr(noc$precip_sc, "scaled:scale")

## Model objects:
m_on <- readRDS(here(dir_305_models, "305c_mod_on_50.rds"))
m_ev <- readRDS(here(dir_305_models, "305d_mod_ev_50.rds"))
m_noc <- readRDS(here(dir_305_models, "305j_mod_noc_50.rds"))
```

```{r}
#| label: palette
color_palette <- c(
  "Onset" = "#009E73",
  "Cessation" = "#D55E00",
  "Low" = "#2B5C8F", # Low Light pollution
  "High" = "#E69F00", # High light
  "Clouds" = "#8997af",
  "No Clouds" = "#272526"
)
```

# Prepare Prediction Data

```{r}
#| label: prep-onset-predictions
# ==========================================
# 1. DIURNAL ONSET (Family-Level Variability)
# ==========================================

alan_seq_on <- seq(
  min(diurn_on$alan_sc, na.rm = TRUE),
  max(diurn_on$alan_sc, na.rm = TRUE),
  length.out = 200
)

cloud_vals_on <- c(
  min(diurn_on$cld_sc, na.rm = TRUE),
  max(diurn_on$cld_sc, na.rm = TRUE)
)

pred_fam_on <- expand.grid(
  alan_sc = alan_seq_on,
  cld_sc = cloud_vals_on,
  precip_sc = precip_0_on,
  family = unique(diurn_on$family),
  sp_grid_wk = unique(diurn_on$sp_grid_wk)[1]
)

pred_fam_on$family <- factor(
  pred_fam_on$family,
  levels = unique(diurn_on$family)
)
pred_fam_on$sp_grid_wk <- factor(
  pred_fam_on$sp_grid_wk,
  levels = unique(diurn_on$sp_grid_wk)
)

# pred_on_raw <- predict(
#   m_on,
#   newdata = pred_fam_on,
#   type = "link",
#   re.form = NULL,
#   allow.new.levels = TRUE,
#   se.fit = TRUE
# )

### Save the prediction objects for later because they take a long time to predict
#saveRDS(pred_on_raw, file = here(dir_307_fig4, "307a_pred_diurn_on.rds"))
### Read in the prediction objects
pred_on_raw <- readRDS(here(dir_307_fig4, "307a_pred_diurn_on.rds"))

# Identity link (Gaussian) -> No back-transformation needed
pred_fam_on$fit <- pred_on_raw$fit
pred_fam_on$lwr <- pred_on_raw$fit - 1.96 * pred_on_raw$se.fit
pred_fam_on$upr <- pred_on_raw$fit + 1.96 * pred_on_raw$se.fit

pred_fam_on <- pred_fam_on |>
  mutate(
    cloud_cat = ifelse(cld_sc == min(cld_sc), "No Clouds", "Clouds"),
    alan = alan_sc *
      sd(diurn_on$alan, na.rm = TRUE) +
      mean(diurn_on$alan, na.rm = TRUE)
  )
```


```{r}
#| label: prep-cess-predictions
# ==========================================
# 2. DIURNAL CESSATION (Family-Level Variability)
# ==========================================

alan_seq_ev <- seq(
  min(diurn_ev$alan_sc, na.rm = TRUE),
  max(diurn_ev$alan_sc, na.rm = TRUE),
  length.out = 200
)

cloud_vals_ev <- c(
  min(diurn_ev$cld_sc, na.rm = TRUE),
  max(diurn_ev$cld_sc, na.rm = TRUE)
)

pred_fam_ev <- expand.grid(
  alan_sc = alan_seq_ev,
  cld_sc = cloud_vals_ev,
  precip_sc = precip_0_ev,
  family = unique(diurn_ev$family),
  sp_grid_wk = unique(diurn_ev$sp_grid_wk)[1]
)

pred_fam_ev$family <- factor(
  pred_fam_ev$family,
  levels = unique(diurn_ev$family)
)
pred_fam_ev$sp_grid_wk <- factor(
  pred_fam_ev$sp_grid_wk,
  levels = unique(diurn_ev$sp_grid_wk)
)

# pred_ev_raw <- predict(
#   m_ev,
#   newdata = pred_fam_ev,
#   type = "link",
#   re.form = NULL,
#   allow.new.levels = TRUE,
#   se.fit = TRUE
# )

### Save the prediction objects for later because they take a long time to predict
#saveRDS(pred_ev_raw, file = here(dir_307_fig4, "307b_pred_diurn_ev.rds"))
### Read in the prediction objects
pred_ev_raw <- readRDS(here(dir_307_fig4, "307b_pred_diurn_ev.rds"))

# Identity link (Gaussian) -> No back-transformation needed
pred_fam_ev$fit <- pred_ev_raw$fit
pred_fam_ev$lwr <- pred_ev_raw$fit - 1.96 * pred_ev_raw$se.fit
pred_fam_ev$upr <- pred_ev_raw$fit + 1.96 * pred_ev_raw$se.fit

pred_fam_ev <- pred_fam_ev |>
  mutate(
    cloud_cat = ifelse(cld_sc == min(cld_sc), "No Clouds", "Clouds"),
    alan = alan_sc *
      sd(diurn_ev$alan, na.rm = TRUE) +
      mean(diurn_ev$alan, na.rm = TRUE)
  )
```

```{r}
#| label: prep-noc-predictions
# ==========================================
# 3. FOR NOCTURNAL (Species-Level Variability)
# ==========================================

alan_seq_noc <- seq(
  min(noc$alan_sc, na.rm = TRUE),
  max(noc$alan_sc, na.rm = TRUE),
  length.out = 200
)

cloud_vals_noc <- c(
  min(noc$cloud_sc, na.rm = TRUE),
  max(noc$cloud_sc, na.rm = TRUE)
)

pred_species <- expand.grid(
  alan_sc = alan_seq_noc,
  cloud_sc = cloud_vals_noc,
  precip_sc = precip_0_noc,
  common_name = unique(noc$common_name),
  sp_grid_wk = unique(noc$sp_grid_wk)[1]
)

pred_species$common_name <- factor(
  pred_species$common_name,
  levels = unique(noc$common_name)
)
pred_species$sp_grid_wk <- factor(
  pred_species$sp_grid_wk,
  levels = unique(noc$sp_grid_wk)
)

# Predict
# pred_noc_raw <- predict(
#   m_noc,
#   newdata = pred_species,
#   type = "link",
#   re.form = NULL,
#   allow.new.levels = TRUE,
#   se.fit = TRUE
# )

### Save the prediction objects for later because they take a long time to predict
#saveRDS(pred_noc_raw, file = here(dir_307_fig4, "307c_pred_noc.rds"))
### Read in the prediction objects
pred_noc_raw <- readRDS(here(dir_307_fig4, "307c_pred_noc.rds"))

pred_species$fit <- plogis(pred_noc_raw$fit)
pred_species$lwr <- plogis(pred_noc_raw$fit - 1.96 * pred_noc_raw$se.fit)
pred_species$upr <- plogis(pred_noc_raw$fit + 1.96 * pred_noc_raw$se.fit)

pred_species <- pred_species |>
  mutate(
    cloud_cat = ifelse(cloud_sc == min(cloud_sc), "No Clouds", "Clouds"),
    alan = alan_sc * sd(noc$alan, na.rm = TRUE) + mean(noc$alan, na.rm = TRUE)
  )
```

# Script 307 - Visualize All Variability

```{r}
#| label: viz-all-onset-fam-var
plot_diurn_on <- ggplot(
  pred_fam_on,
  aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  #facet_wrap(~family, scales = "free_y") + # uncomment for freey
  facet_wrap(~family) + # comment for freey
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  coord_cartesian(ylim = c(-4.0, 4.0)) + # comment for freey
  labs(
    title = "Diurnal Onset - Family Variability",
    x = "Light Pollution",
    y = "Vocal Onset\n(hours relative to sunrise)",
    color = "Cloud cover",
    fill = "Cloud cover"
  ) +
  theme_classic(base_size = 18) +
  theme(
    legend.position = "bottom"
  )


ggsave(
  plot_diurn_on,
  filename = here(dir_307_fig4, "307_FigS1.png"),
  #filename = here(dir_307_fig4, "307_FigS1_freey.png"),
  device = "png",
  height = 12,
  width = 16,
  units = "in",
  dpi = 600
)
```

```{r}
#| label: viz-all-cess-fam-var
plot_diurn_ev <- ggplot(
  pred_fam_ev,
  aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  #facet_wrap(~family, scales = "free_y") + # uncomment for freey
  facet_wrap(~family) + # comment for freey
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  coord_cartesian(ylim = c(-4.0, 4.0)) + # comment for freey
  labs(
    title = "Diurnal Cessation - Family Variability",
    x = "Light Pollution",
    y = "Vocal Cessation\n(hours relative to sunset)",
    color = "Cloud cover",
    fill = "Cloud cover"
  ) +
  theme_classic(base_size = 18) +
  theme(
    legend.position = "bottom"
  )


ggsave(
  plot_diurn_ev,
  filename = here(dir_307_fig4, "307_FigS2.png"),
  #filename = here(dir_307_fig4, "307_FigS2_freey.png"),
  device = "png",
  height = 12,
  width = 16,
  units = "in",
  dpi = 600
)
```


```{r}
#| label: viz-all-noc-spp-var
plot_noc_sp <- ggplot(
  pred_species,
  aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  facet_wrap(~common_name) +
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  labs(
    title = "Nocturnal - All Species Variability",
    x = "Light Pollution",
    y = "Probability of Vocalization",
    color = "Cloud cover",
    fill = "Cloud cover"
  ) +
  theme_classic(base_size = 18) +
  theme(
    legend.position = "bottom"
  )


ggsave(
  plot_noc_sp,
  filename = here(dir_307_fig4, "307_FigS3.png"),
  device = "png",
  height = 12,
  width = 16,
  units = "in",
  dpi = 600
)
```

# Script 307 Extract Coefficients

```{r}
#| label: fx-extract-alan-coefs
extract_alan_coefficients <- function(model_obj, model_name) {
  # Get global fixed effect for ALAN
  fixed_effects <- fixef(model_obj)$cond
  alan_fixed <- fixed_effects["alan_sc"]

  # Get the random effect deviations for each family
  random_effects <- ranef(model_obj)$cond$family

  # Calculate total family-level beta coefficient
  coef_table <- random_effects |>
    rownames_to_column(var = "family") |>
    rename(random_deviation = alan_sc) |>
    mutate(
      model = model_name,
      fixed_baseline = alan_fixed,
      total_alan_beta = fixed_baseline + random_deviation
    ) |>
    select(model, family, fixed_baseline, random_deviation, total_alan_beta)

  return(coef_table)
}
```

```{r}
#| label: create-coef-md-table
# Extract for both diurnal models
coef_on <- extract_alan_coefficients(m_on, "on")
coef_ev <- extract_alan_coefficients(m_ev, "ev")

# Grab number of vocs data for each family
fam_num_on <- diurn_on |>
  group_by(family) |>
  summarize(num_vocs_on = n(), .groups = "drop")

fam_num_ev <- diurn_ev |>
  group_by(family) |>
  summarize(num_vocs_ev = n(), .groups = "drop")


# Combine into a single master data frame
alan_beta_table <- bind_rows(coef_on, coef_ev)
print(alan_beta_table)


alan_beta_wide <- alan_beta_table |>
  select(family, model, total_alan_beta) |>
  # Pivot from long to wide format
  pivot_wider(
    names_from = model,
    values_from = total_alan_beta,
    names_prefix = "total_alan_beta_"
  ) |>
  left_join(fam_num_on, by = "family") |>
  left_join(fam_num_ev, by = "family")
print(alan_beta_wide, n = 63)

kable(
  alan_beta_wide,
  digits = 3,
  format = "markdown",
  col.names = c(
    "Family",
    "Total Beta (Onset)",
    "Total Beta (Cess)",
    "Num. Vocs (Onset)",
    "Num. Vocs (Cess)"
  )
)
```


# Create Figure 4/5 Panels

```{r}
#| label: data-prep
# --- Extract Diurnal Family Coefficients (Onset & Cessation)
coef_on <- extract_alan_coefficients(m_on, "Onset")
coef_ev <- extract_alan_coefficients(m_ev, "Cessation")

# Get sample size per family (vocs count)
fam_num_on <- diurn_on |>
  group_by(family) |>
  summarize(n_vocs_on = n(), .groups = "drop")

fam_num_ev <- diurn_ev |>
  group_by(family) |>
  summarize(n_vocs_ev = n(), .groups = "drop")

# Combine diurnal coefficients for Panel A
coef_panel_a <- bind_rows(coef_on, coef_ev) |>
  left_join(fam_num_on, by = "family") |>
  left_join(fam_num_ev, by = "family") |>
  mutate(
    n_vocs = ifelse(model == "Onset", n_vocs_on, n_vocs_ev)
  )

# --- Extract Nocturnal Species Coefficients
# Define Dark (5th percentile) and Bright (95th percentile) ALAN values
alan_low <- quantile(noc$alan_sc, probs = 0.05, na.rm = TRUE)
alan_high <- quantile(noc$alan_sc, probs = 0.95, na.rm = TRUE)

# Build grid for predictions (hold cloud_sc at 0 = mean cloud cover)
grid_noc <- expand.grid(
  alan_sc = c(alan_low, alan_high),
  cloud_sc = 0, # Mean cloud cover baseline
  precip_sc = precip_0_noc,
  common_name = unique(noc$common_name),
  sp_grid_wk = unique(noc$sp_grid_wk)[1]
)

# Predict link scale (log-odds) and transform to probability (0-1)
preds_noc_raw <- predict(
  m_noc,
  newdata = grid_noc,
  type = "link",
  re.form = NULL,
  allow.new.levels = TRUE
)

grid_noc <- grid_noc |>
  mutate(
    prob_det = plogis(preds_noc_raw),
    light_condition = ifelse(
      alan_sc == alan_low,
      "Low",
      "High"
    )
  )

# Noc species vocalization count
sp_num_noc <- noc |>
  group_by(common_name) |>
  summarize(n_vocs = n(), .groups = "drop")
```

## Fig. 4 - Diurnal family & species

```{r}
#| label: create-fig4-panel-A
# Panel A: Family-level ALAN Slopes (Ordered by Onset Coefficients)

# Create wide format table and arrange families by Onset coefficient
family_order_on <- coef_panel_a |>
  select(family, model, total_alan_beta) |>
  pivot_wider(names_from = model, values_from = total_alan_beta) |>
  arrange(Onset)

# Apply ordered factor sequence back to main data
coef_panel_a_on <- coef_panel_a |>
  mutate(
    family = factor(family, levels = family_order_on$family)
  )

coef_panel_a_on_wide <- family_order_on |>
  mutate(family = factor(family, levels = family_order_on$family))

# Create numeric Y-index for background row shading
n_fams <- nrow(family_order_on)
zebra_df <- data.frame(
  ymin = seq(0.5, n_fams - 0.5, by = 2),
  ymax = seq(1.5, n_fams + 0.5, by = 2)
)

# Create Panel A
plot_panel_a <- ggplot() +
  # Row shading
  geom_rect(
    data = zebra_df,
    aes(xmin = -Inf, xmax = Inf, ymin = ymin, ymax = ymax),
    fill = "grey96",
    color = NA
  ) +
  # Zero reference line
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    color = "grey60",
    alpha = 0.7
  ) +
  # Dumbbell segment connecting Onset and Cessation
  geom_segment(
    data = coef_panel_a_on_wide,
    aes(y = family, yend = family, x = Onset, xend = Cessation),
    color = "grey70",
    linewidth = 0.5,
    na.rm = TRUE
  ) +
  # Coefficient dots
  geom_point(
    data = coef_panel_a_on,
    aes(x = total_alan_beta, y = family, color = model, size = n_vocs),
    alpha = 0.85
  ) +
  scale_color_manual(values = color_palette) +
  #scale_fill_manual(values = color_palette) +
  labs(
    title = "A) Diurnal Family-Level\nLight Pollution Slopes",
    x = expression(beta ~ "Coefficient (Light Pollution Response)"),
    y = "Family",
    color = "Response"
  ) +
  # Continuous size scale
  scale_size(
    range = c(1, 5),
    name = "Number of\nVocalizations",
    breaks = c(1000, 5000, 15000, 50000),
    labels = scales::label_comma()
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "right",
    axis.text.y = element_text(size = 10, color = "black"),
    panel.grid.major = element_blank()
  )

print(plot_panel_a)
```


```{r}
#| label: create-fig4-panel-B-HORIZONTAL
# Panel B: Selected Family Responses to ALAN (Onset & Cessation)

# Specify families based on expected, neutral, and opposite responses to ALAN
selected_fams <- c(
  "Corvidae", # expected
  "Sittidae", # neutral
  "Cuculidae" # opposite
)

# Sub-label for facet ordering
fam_labels <- c(
  "Corvidae" = "Expected Relationship",
  "Sittidae" = "Neutral Relationship",
  "Cuculidae" = "Opposite Relationship"
)


# Filter predictions for selected families
pred_sub_on <- pred_fam_on |>
  filter(family %in% selected_fams) |>
  mutate(
    response_type = "Onset",
    rel_type = factor(
      fam_labels[as.character(family)],
      levels = c(
        "Expected Relationship",
        "Neutral Relationship",
        "Opposite Relationship"
      )
    )
  )


pred_sub_ev <- pred_fam_ev |>
  filter(family %in% selected_fams) |>
  mutate(
    response_type = "Cessation",
    rel_type = factor(
      fam_labels[as.character(family)],
      levels = c(
        "Expected Relationship",
        "Neutral Relationship",
        "Opposite Relationship"
      )
    )
  )


# FX: Create Onset/Cessation plots
plot_response_row <- function(data, y_label, y_limits = c(-2.5, 2.5)) {
  ggplot(data, aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)) +
    # Reference line at y = 0
    geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = "grey60",
      linewidth = 0.4
    ) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    facet_wrap(~ rel_type + family, nrow = 1) +
    scale_color_manual(values = color_palette) +
    scale_fill_manual(values = color_palette) +
    scale_y_continuous(
      limits = rev(sort(y_limits)),
      trans = "reverse",
      expand = c(0, 0)
    ) +
    labs(
      x = "Light Pollution",
      y = y_label,
      color = "Cloud Cover",
      fill = "Cloud Cover"
    ) +
    theme_classic(base_size = 18) +
    theme(
      #  axis.title.y = element_text(margin = margin(r = -20)), # change to fix Y axis label
      strip.background = element_rect(fill = "grey95", color = NA),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.3),
      panel.grid.minor = element_line(color = "grey96", linewidth = 0.2),
      plot.margin = margin(t = 5, r = 5, b = 5, l = 5)
    )
}


# FX: Sunrise/sunset guide
build_y_key_panel <- function(
  solar_event = "Sunrise",
  y_limits = c(-2.5, 2.5),
  fixed_arrow_len = 1.2
) {
  y_min <- min(y_limits)
  y_max <- max(y_limits)
  y_span <- y_max - y_min

  # Fixed Arrow
  up_start <- 0.3
  up_end <- 0.3 + fixed_arrow_len

  down_start <- -0.3
  down_end <- -0.3 - fixed_arrow_len

  ggplot(data.frame(x = c(0, 1), y = y_limits), aes(x = x, y = y)) +
    # Reference line at y = 0
    geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = "grey60",
      linewidth = 0.4
    ) +

    # Sunrise/Sunset Label
    annotate(
      "text",
      x = 0.05,
      y = y_span * 0.015,
      label = solar_event,
      hjust = 0,
      vjust = -0.2,
      size = 4,
      fontface = "bold",
      color = "grey20"
    ) +

    # Fixed Upward Arrow: Hours After
    annotate(
      "segment",
      x = 0.2,
      xend = 0.2,
      y = up_start,
      yend = up_end,
      arrow = arrow(length = unit(0.12, "cm")),
      color = "grey40",
      linewidth = 0.5
    ) +
    annotate(
      "text",
      x = 0.32,
      y = (up_start + up_end) / 2,
      label = paste("Hours After\n", solar_event),
      hjust = 0,
      size = 4,
      fontface = "italic",
      color = "grey30"
    ) +

    # Fixed Downward Arrow: Hours Before
    annotate(
      "segment",
      x = 0.2,
      xend = 0.2,
      y = down_start,
      yend = down_end,
      arrow = arrow(length = unit(0.12, "cm")),
      color = "grey40",
      linewidth = 0.5
    ) +
    annotate(
      "text",
      x = 0.32,
      y = (down_start + down_end) / 2,
      label = paste("Hours Before\n", solar_event),
      hjust = 0,
      size = 4,
      fontface = "italic",
      color = "grey30"
    ) +

    scale_y_continuous(
      limits = c(y_min, y_max),
      trans = "reverse",
      expand = c(0, 0)
    ) +
    scale_x_continuous(limits = c(0, 1.2), expand = c(0, 0)) +
    labs(title = NULL, y = NULL, x = NULL) +
    theme_void(base_size = 18) +
    theme(
      plot.title = element_text(
        size = 8.5,
        face = "bold",
        color = "grey30",
        hjust = 0.5,
        margin = margin(b = 5)
      ),
      plot.margin = margin(t = 5, r = 2, b = 5, l = 8)
    )
}


# Plot top row (diurnal onset)
onset_limits <- c(-0.5, 5.0)

plot_families_on <- plot_response_row(
  data = pred_sub_on,
  y_label = "Vocal Onset\n(hours relative to sunrise)",
  y_limits = onset_limits
) +
  labs(
    title = "B) Diurnal Family-Level Predicted Responses to\nLight Pollution and Cloud Cover"
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

plot_key_on <- build_y_key_panel(
  solar_event = "Sunrise",
  y_limits = onset_limits
)

panel_b_top <- (plot_families_on | plot_key_on) +
  plot_layout(widths = c(6, 1))
panel_b_top


# Plot bottom row (diurnal cessation)
cessation_limits <- c(-5.0, 0.5)

plot_families_ev <- plot_response_row(
  data = pred_sub_ev,
  y_label = "Vocal Cessation\n(hours relative to sunset)",
  y_limits = cessation_limits
) +
  theme(
    strip.text = element_blank(),
    strip.background = element_blank()
  )

plot_key_ev <- build_y_key_panel(
  solar_event = "Sunset",
  y_limits = cessation_limits
)

panel_b_bottom <- (plot_families_ev | plot_key_ev) +
  plot_layout(widths = c(6, 1))
panel_b_bottom


# Combine to make Panel B ---
plot_panel_b_hori <- (panel_b_top / panel_b_bottom) +
  plot_layout(heights = c(1, 1), guides = "collect") &
  theme(legend.position = "bottom")
plot_panel_b_hori
```


```{r}
#| label: create-fig4-panel-B-VERTICAL
# Panel B: Selected Family Responses to ALAN (Onset & Cessation) - 3 Rows x 2 Columns
selected_fams <- c(
  "Corvidae", # expected
  "Sittidae", # neutral
  "Cuculidae" # opposite
)

fam_labels <- c(
  "Corvidae" = "Expected Relationship: Corvidae",
  "Sittidae" = "Neutral Relationship: Sittidae",
  "Cuculidae" = "Opposite Relationship: Cuculidae"
)

# FX: Gray banner bar spanning the row
build_banner_panel <- function(text_label) {
  ggplot() +
    annotate(
      "rect",
      xmin = -Inf,
      xmax = Inf,
      ymin = -Inf,
      ymax = Inf,
      fill = "grey92",
      color = NA
    ) +
    annotate(
      "text",
      x = 0.5,
      y = 0.5,
      label = text_label,
      size = 4.2,
      fontface = "bold",
      color = "grey20"
    ) +
    theme_void() +
    theme(
      plot.margin = margin(t = 6, r = 6, b = 2, l = 6)
    )
}

# Limits
onset_limits <- c(-0.5, 5.0)
cessation_limits <- c(-5.0, 0.5)

# FX: Single family plot panel
plot_single_panel <- function(
  data,
  f_name,
  y_limits,
  col_title = NULL,
  y_axis_label = NULL,
  show_x_axis = FALSE
) {
  p <- data |>
    filter(family == f_name) |>
    ggplot(aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)) +
    geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = "grey60",
      linewidth = 0.4
    ) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    scale_color_manual(values = color_palette) +
    scale_fill_manual(values = color_palette) +
    scale_x_continuous(
      limits = c(-2, 4),
      breaks = seq(-2, 4, by = 2),
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      limits = rev(sort(y_limits)),
      trans = "reverse",
      expand = c(0, 0)
    ) +
    labs(
      title = col_title,
      x = if (show_x_axis) "Light Pollution: ln(Radiance + 1)" else NULL,
      y = y_axis_label,
      color = "Cloud Cover",
      fill = "Cloud Cover"
    ) +
    theme_classic(base_size = 14) +
    theme(
      plot.title = element_text(
        face = "bold",
        size = 13,
        hjust = 0.5,
        margin = margin(b = 4)
      ),
      axis.title.y = element_text(size = 10, margin = margin(r = 5)),
      axis.title.x = element_text(size = 11, margin = margin(t = 5)),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.3),
      panel.grid.minor = element_line(color = "grey96", linewidth = 0.2),
      plot.margin = margin(t = 4, r = 6, b = 4, l = 6)
    )

  if (!show_x_axis) {
    p <- p +
      theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
      )
  }

  return(p)
}

# FX: Sunrise/sunset guide panel
build_y_key_panel <- function(
  solar_event = "Sunrise",
  y_limits = c(-2.5, 2.5),
  fixed_arrow_len = 1.2
) {
  y_min <- min(y_limits)
  y_max <- max(y_limits)
  y_span <- y_max - y_min

  up_start <- 0.3
  up_end <- 0.3 + fixed_arrow_len

  down_start <- -0.3
  down_end <- -0.3 - fixed_arrow_len

  ggplot(data.frame(x = c(0, 1), y = y_limits), aes(x = x, y = y)) +
    geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = "grey60",
      linewidth = 0.4
    ) +
    annotate(
      "text",
      x = 0.05,
      y = y_span * 0.015,
      label = solar_event,
      hjust = 0,
      vjust = -0.2,
      size = 3.2,
      fontface = "bold",
      color = "grey20"
    ) +
    annotate(
      "segment",
      x = 0.2,
      xend = 0.2,
      y = up_start,
      yend = up_end,
      arrow = arrow(length = unit(0.1, "cm")),
      color = "grey40",
      linewidth = 0.4
    ) +
    annotate(
      "text",
      x = 0.32,
      y = (up_start + up_end) / 2,
      label = paste("Hours After\n", solar_event),
      hjust = 0,
      size = 2.8,
      fontface = "italic",
      color = "grey30"
    ) +
    annotate(
      "segment",
      x = 0.2,
      xend = 0.2,
      y = down_start,
      yend = down_end,
      arrow = arrow(length = unit(0.1, "cm")),
      color = "grey40",
      linewidth = 0.4
    ) +
    annotate(
      "text",
      x = 0.32,
      y = (down_start + down_end) / 2,
      label = paste("Hours Before\n", solar_event),
      hjust = 0,
      size = 2.8,
      fontface = "italic",
      color = "grey30"
    ) +
    scale_y_continuous(
      limits = c(y_min, y_max),
      trans = "reverse",
      expand = c(0, 0)
    ) +
    scale_x_continuous(limits = c(0, 1.2), expand = c(0, 0)) +
    labs(title = NULL, y = NULL, x = NULL) +
    theme_void() +
    theme(plot.margin = margin(t = 15, r = 2, b = 5, l = 2))
}

# Static key panels for each column
key_on <- build_y_key_panel("Sunrise", onset_limits)
key_ev <- build_y_key_panel("Sunset", cessation_limits)

onset_txt <- "Vocal Onset\n(hours relative to sunrise)"
cess_txt <- "Vocal Cessation\n(hours relative to sunset)"

# Build all panels as individual plot objects
b1 <- build_banner_panel(fam_labels["Corvidae"])
p1_on <- plot_single_panel(
  pred_fam_on,
  "Corvidae",
  onset_limits,
  y_axis_label = onset_txt,
  show_x_axis = FALSE
)
p1_ev <- plot_single_panel(
  pred_fam_ev,
  "Corvidae",
  cessation_limits,
  y_axis_label = cess_txt,
  show_x_axis = FALSE
)

b2 <- build_banner_panel(fam_labels["Sittidae"])
p2_on <- plot_single_panel(
  pred_fam_on,
  "Sittidae",
  onset_limits,
  y_axis_label = onset_txt,
  show_x_axis = FALSE
)
p2_ev <- plot_single_panel(
  pred_fam_ev,
  "Sittidae",
  cessation_limits,
  y_axis_label = cess_txt,
  show_x_axis = FALSE
)

b3 <- build_banner_panel(fam_labels["Cuculidae"])
p3_on <- plot_single_panel(
  pred_fam_on,
  "Cuculidae",
  onset_limits,
  y_axis_label = onset_txt,
  show_x_axis = TRUE
)
p3_ev <- plot_single_panel(
  pred_fam_ev,
  "Cuculidae",
  cessation_limits,
  y_axis_label = cess_txt,
  show_x_axis = TRUE
)

# Layout design matrix
design <- "
  AAAA
  BCDE
  FFFF
  GHIJ
  KKKK
  LMNO
"

plot_panel_b_vert <- wrap_plots(
  A = b1,
  B = p1_on,
  C = key_on,
  D = p1_ev,
  E = key_ev,
  F = b2,
  G = p2_on,
  H = key_on,
  I = p2_ev,
  J = key_ev,
  K = b3,
  L = p3_on,
  M = key_on,
  N = p3_ev,
  O = key_ev,
  design = design,
  widths = c(4, 1, 4, 1),
  heights = c(0.18, 1, 0.18, 1, 0.18, 1),
  guides = "collect"
) &
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11)
  )

plot_panel_b_vert
```

```{r}
#| label: create-fig4-panel-B-NOKEY
# Panel B: Selected Family Responses to ALAN (Onset & Cessation) - 3 Rows x 2 Columns
selected_fams <- c(
  "Corvidae", # expected
  "Sittidae", # neutral
  "Cuculidae" # opposite
)

fam_labels <- c(
  "Corvidae" = "Expected Relationship: Corvidae",
  "Sittidae" = "Neutral Relationship: Sittidae",
  "Cuculidae" = "Opposite Relationship: Cuculidae"
)

# FX: Gray banner bar spanning the row
build_banner_panel <- function(text_label) {
  ggplot() +
    annotate(
      "rect",
      xmin = -Inf,
      xmax = Inf,
      ymin = -Inf,
      ymax = Inf,
      fill = "grey92",
      color = NA
    ) +
    annotate(
      "text",
      x = 0.5,
      y = 0.5,
      label = text_label,
      size = 4.2,
      fontface = "bold",
      color = "grey20"
    ) +
    theme_void() +
    theme(
      plot.margin = margin(t = 6, r = 6, b = 2, l = 6)
    )
}

# Limits
onset_limits <- c(-0.5, 5.0)
cessation_limits <- c(-5.0, 0.5)

# FX: Single family plot panel
plot_single_panel <- function(
  data,
  f_name,
  y_limits,
  col_title = NULL,
  y_axis_label = NULL,
  show_x_axis = FALSE
) {
  p <- data |>
    filter(family == f_name) |>
    ggplot(aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)) +
    geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = "grey60",
      linewidth = 0.4
    ) +
    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    scale_color_manual(values = color_palette) +
    scale_fill_manual(values = color_palette) +
    scale_x_continuous(
      limits = c(-2, 4),
      breaks = seq(-2, 4, by = 2),
      expand = c(0, 0)
    ) +
    scale_y_continuous(
      limits = rev(sort(y_limits)),
      trans = "reverse",
      expand = c(0, 0)
    ) +
    labs(
      title = col_title,
      x = if (show_x_axis) "Light Pollution: ln(Radiance + 1)" else NULL,
      y = y_axis_label,
      color = "Cloud Cover",
      fill = "Cloud Cover"
    ) +
    theme_classic(base_size = 14) +
    theme(
      plot.title = element_text(
        face = "bold",
        size = 13,
        hjust = 0.5,
        margin = margin(b = 4)
      ),
      axis.title.y = element_text(size = 10, margin = margin(r = 5)),
      axis.title.x = element_text(size = 11, margin = margin(t = 5)),
      panel.grid.major = element_line(color = "grey92", linewidth = 0.3),
      panel.grid.minor = element_line(color = "grey96", linewidth = 0.2),
      plot.margin = margin(t = 4, r = 6, b = 4, l = 6)
    )

  if (!show_x_axis) {
    p <- p +
      theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
      )
  }

  return(p)
}

# FX: Sunrise/sunset guide panel
build_y_key_panel <- function(
  solar_event = "Sunrise",
  y_limits = c(-2.5, 2.5),
  fixed_arrow_len = 1.2
) {
  y_min <- min(y_limits)
  y_max <- max(y_limits)
  y_span <- y_max - y_min

  up_start <- 0.3
  up_end <- 0.3 + fixed_arrow_len

  down_start <- -0.3
  down_end <- -0.3 - fixed_arrow_len

  ggplot(data.frame(x = c(0, 1), y = y_limits), aes(x = x, y = y)) +
    geom_hline(
      yintercept = 0,
      linetype = "dashed",
      color = "grey60",
      linewidth = 0.4
    ) +
    annotate(
      "text",
      x = 0.05,
      y = y_span * 0.015,
      label = solar_event,
      hjust = 0,
      vjust = -0.2,
      size = 3.2,
      fontface = "bold",
      color = "grey20"
    ) +
    annotate(
      "segment",
      x = 0.2,
      xend = 0.2,
      y = up_start,
      yend = up_end,
      arrow = arrow(length = unit(0.1, "cm")),
      color = "grey40",
      linewidth = 0.4
    ) +
    annotate(
      "text",
      x = 0.32,
      y = (up_start + up_end) / 2,
      label = paste("Hours After\n", solar_event),
      hjust = 0,
      size = 2.8,
      fontface = "italic",
      color = "grey30"
    ) +
    annotate(
      "segment",
      x = 0.2,
      xend = 0.2,
      y = down_start,
      yend = down_end,
      arrow = arrow(length = unit(0.1, "cm")),
      color = "grey40",
      linewidth = 0.4
    ) +
    annotate(
      "text",
      x = 0.32,
      y = (down_start + down_end) / 2,
      label = paste("Hours Before\n", solar_event),
      hjust = 0,
      size = 2.8,
      fontface = "italic",
      color = "grey30"
    ) +
    scale_y_continuous(
      limits = c(y_min, y_max),
      trans = "reverse",
      expand = c(0, 0)
    ) +
    scale_x_continuous(limits = c(0, 1.2), expand = c(0, 0)) +
    labs(title = NULL, y = NULL, x = NULL) +
    theme_void() +
    theme(plot.margin = margin(t = 15, r = 2, b = 5, l = 2))
}

# Static key panels for each column
key_on <- build_y_key_panel("Sunrise", onset_limits)
key_ev <- build_y_key_panel("Sunset", cessation_limits)

onset_txt <- "Vocal Onset\n(hours relative to sunrise)"
cess_txt <- "Vocal Cessation\n(hours relative to sunset)"

# Build all panels as individual plot objects
b1 <- build_banner_panel(fam_labels["Corvidae"])
p1_on <- plot_single_panel(
  pred_fam_on,
  "Corvidae",
  onset_limits,
  y_axis_label = onset_txt,
  show_x_axis = FALSE
)
p1_ev <- plot_single_panel(
  pred_fam_ev,
  "Corvidae",
  cessation_limits,
  y_axis_label = cess_txt,
  show_x_axis = FALSE
)

b2 <- build_banner_panel(fam_labels["Sittidae"])
p2_on <- plot_single_panel(
  pred_fam_on,
  "Sittidae",
  onset_limits,
  y_axis_label = onset_txt,
  show_x_axis = FALSE
)
p2_ev <- plot_single_panel(
  pred_fam_ev,
  "Sittidae",
  cessation_limits,
  y_axis_label = cess_txt,
  show_x_axis = FALSE
)

b3 <- build_banner_panel(fam_labels["Cuculidae"])
p3_on <- plot_single_panel(
  pred_fam_on,
  "Cuculidae",
  onset_limits,
  y_axis_label = onset_txt,
  show_x_axis = TRUE
) +
  theme(axis.title.x = element_blank())

p3_ev <- plot_single_panel(
  pred_fam_ev,
  "Cuculidae",
  cessation_limits,
  y_axis_label = cess_txt,
  show_x_axis = TRUE
) +
  theme(axis.title.x = element_blank())

x_label_panel <- ggplot() +
  annotate(
    "text",
    x = 0.5,
    y = 0.5,
    label = "Light Pollution: ln(Radiance + 1)",
    size = 4,
    fontface = "plain"
  ) +
  theme_void() +
  theme(plot.margin = margin(t = 2, b = 2))

# Layout design matrix
design <- "
  AA
  BC
  DD
  EF
  GG
  HI
  JJ
"

plot_panel_b_nokey <- wrap_plots(
  A = b1,
  B = p1_on,
  C = p1_ev,
  D = b2,
  E = p2_on,
  F = p2_ev,
  G = b3,
  H = p3_on,
  I = p3_ev,
  J = x_label_panel,
  design = design,
  widths = c(4, 4),
  heights = c(0.18, 1),
  guides = "collect"
) &
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 11)
  )

plot_panel_b_nokey
```

```{r}
#| label: select-panel-b
##plot_panel_b_final <- plot_panel_b_hori
##plot_panel_b_final <- plot_panel_b_vert
plot_panel_b_final <- plot_panel_b_nokey # selected for final visualization

# add plot title
plot_panel_b_final <- plot_panel_b_final +
  plot_annotation(
    title = "B) Selected Diurnal Family Responses\nto Light Pollution and Cloud Cover",
    theme = theme(
      plot.title = element_text(size = 16, hjust = 0, margin = margin(b = 10))
    )
  )
```

```{r}
#| label: save-fig4-individually
# Save individual panels
# Panel A - Diurnal Family Coefficients
ggsave(
  filename = here(dir_307_fig4, "307_fig4_panelA.png"),
  plot = plot_panel_a,
  device = "png",
  height = 9,
  width = 9,
  units = "in",
  dpi = 300
)

# Panel B - Family Predicted ALAN Response
ggsave(
  filename = here(dir_307_fig4, "307_fig4_panelB.png"),
  plot = plot_panel_b_final,
  device = "png",
  height = 9,
  width = 9,
  units = "in",
  dpi = 300
)
```

```{r}
#| label: save-fig4-combined
# Panels A & B side-by-side
final_fig4 <- (wrap_elements(plot_panel_a) |
  wrap_elements(plot_panel_b_final)) +
  plot_layout(widths = c(1.3, 1))
final_fig4

# Save layout
ggsave(
  filename = here(dir_307_fig4, "307_fig4.png"),
  plot = final_fig4,
  device = "png",
  height = 10,
  width = 12,
  units = "in",
  dpi = 300
)
```


## Fig. 5 - Nocturnal Families

Grab a couple intesting nocturnal families

Three owls:
Eastern Screech owl - expected
Tawny owl - neutral
Tropical screech-owl -opposite

Three nightjars:
Australian owlet-nightjar - expected
eurasian nightjar/common poorwill - neutral
Eastern Whippoorwill - opposite


```{r}
#| label: create-fig5-panel-A
# Panel A: Dark vs Bright Light Pollution probabilities (Ordered by dark levels)

# Setup data in long format
coef_panel_a_long <- grid_noc |>
  left_join(sp_num_noc, by = "common_name")

# Setup data in wide format - order by dark conditions
coef_panel_a_wide <- grid_noc |>
  select(common_name, light_condition, prob_det) |>
  pivot_wider(names_from = light_condition, values_from = prob_det) |>
  rename(Dark = `Low`, Bright = `High`) |>
  # Arrange species continuously from lowest detection in the dark to highest
  arrange(Dark)

# Re-level common_name factor
coef_panel_a_long <- coef_panel_a_long |>
  mutate(
    common_name = factor(common_name, levels = coef_panel_a_wide$common_name)
  )

coef_panel_a_wide <- coef_panel_a_wide |>
  mutate(
    common_name = factor(common_name, levels = coef_panel_a_wide$common_name)
  )

# Setup background grey/white row shading
n_species <- nrow(coef_panel_a_wide)
zebra_df_b <- data.frame(
  ymin = seq(0.5, n_species - 0.5, by = 2),
  ymax = seq(1.5, n_species + 0.5, by = 2)
)

# Plot Panel A
plot_fig5_panel_a <- ggplot() +
  # Row shading
  geom_rect(
    data = zebra_df_b,
    aes(xmin = -Inf, xmax = Inf, ymin = ymin, ymax = ymax),
    fill = "grey96",
    color = NA
  ) +
  # Dumbbell segment connecting Dark and Bright predictions
  geom_segment(
    data = coef_panel_a_wide,
    aes(y = common_name, yend = common_name, x = Dark, xend = Bright),
    color = "grey70",
    linewidth = 0.5
  ) +
  # Detection probability points
  geom_point(
    data = coef_panel_a_long,
    aes(x = prob_det, y = common_name, color = light_condition, size = n_vocs),
    alpha = 0.85
  ) +
  # Colors for Light pollution levels
  scale_color_manual(values = color_palette) +
  # Scale size of dots
  scale_size(
    range = c(1, 7.5),
    name = "Number of\nVocalizations",
    breaks = c(1500, 10000, 35000, 100000),
    labels = scales::label_comma()
  ) +
  # Format X-axis as clean %s
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0, 1)
  ) +
  labs(
    title = "A) Nocturnal Species-Level\nPredicted Vocalization Probability",
    x = "Probability of Vocalization",
    y = "Species",
    color = "Level of\nLight Pollution"
  ) +
  theme_classic(base_size = 18) +
  theme(
    legend.position = "right",
    axis.text.y = element_text(size = 12, color = "black"),
    panel.grid.major = element_blank()
  )

print(plot_fig5_panel_a)
```


```{r}
#| label: create-fig5-panel-b
# Panel B: Selected Species Responses to ALAN (Nocturnal)

# Specify species based on expected, neutral, and opposite responses to ALAN
species_rel_map <- c(
  "Eastern Screech-Owl" = "Expected Relationship",
  "Tawny Owl" = "Neutral Relationship",
  "Tropical Screech-Owl" = "Opposite Relationship",
  "Australian Owlet-nightjar" = "Expected Relationship",
  "Eurasian Nightjar" = "Neutral Relationship",
  "Eastern Whip-poor-will" = "Opposite Relationship"
)

rel_levels <- c(
  "Expected Relationship",
  "Neutral Relationship",
  "Opposite Relationship"
)

species_ordered <- c(
  "Eastern Screech-Owl",
  "Australian Owlet-nightjar",
  "Tawny Owl",
  "Eurasian Nightjar",
  "Tropical Screech-Owl",
  "Eastern Whip-poor-will"
)

# Prepare the subset data
pred_sub <- pred_species |>
  filter(common_name %in% species_ordered) |>
  mutate(
    rel_type = factor(
      species_rel_map[as.character(common_name)],
      levels = rel_levels
    ),
    common_name = factor(common_name, levels = species_ordered)
  )

plot_noc_sp <- ggplot(
  pred_sub,
  aes(x = alan, y = fit, color = cloud_cat, fill = cloud_cat)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, color = NA) +
  facet_nested_wrap(
    vars(rel_type, common_name),
    ncol = 2,
    axes = "y"
  ) +
  scale_color_manual(values = color_palette) +
  scale_fill_manual(values = color_palette) +
  labs(
    title = "B) Selected Nocturnal Species Responses\nto Light Pollution and Cloud Cover",
    x = "Light Pollution: ln(Radiance + 1)",
    y = "Probability of Vocalization",
    color = "Cloud cover",
    fill = "Cloud cover"
  ) +
  theme_classic(base_size = 18) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey95", color = NA),
    strip.text = element_text(face = "bold"),
    ggh4x.facet.nestline = element_line(color = "black") # Optional accent line
  )

print(plot_noc_sp)
```

```{r}
#| label: save-fig5-individually
# Panel A - Nocturnal Predicted ALAN Probability
ggsave(
  filename = here(dir_307_fig4, "307_fig5_panelA.png"),
  plot = plot_fig5_panel_a,
  device = "png",
  height = 9,
  width = 9,
  units = "in",
  dpi = 300
)

# Save panel B
ggsave(
  filename = here(dir_307_fig4, "307_fig5_panelB.png"),
  plot = plot_noc_sp,
  device = "png",
  height = 9,
  width = 8,
  units = "in",
  dpi = 300
)
```

```{r}
#| label: save-fig5-combined
final_fig5 <- plot_fig5_panel_a + plot_noc_sp

# Save combined fig5
ggsave(
  filename = here(dir_307_fig4, "307_fig5.png"),
  plot = final_fig5,
  device = "png",
  height = 10,
  width = 18,
  units = "in",
  dpi = 300
)
```

# Supporting text results

Calculate predictions to support results text

Onset advancement by family...
These numbers are slightly smaller than Pease & Gilbert 2025 because these results use 5th and 95th percentiles of ALAN rather than min/max. Also, the preds vary by family
```{r}
#| label: calc-ALAN-cld-preds
# Calculate ALAN quantiles (5th, 95th)
p05_val <- quantile(diurn_on$alan_sc, 0.05, na.rm = TRUE)
p95_val <- quantile(diurn_on$alan_sc, 0.95, na.rm = TRUE)

# Grab closest ALAN value
closest_p05 <- alan_seq_on[which.min(abs(alan_seq_on - p05_val))]
closest_p95 <- alan_seq_on[which.min(abs(alan_seq_on - p95_val))]

# Calculate onset advancement in minutes per family
family_advances <- pred_fam_on |>
  filter(alan_sc %in% c(closest_p05, closest_p95)) |>
  mutate(alan_cat = ifelse(alan_sc == closest_p05, "Dark", "Light")) |>
  select(family, cloud_cat, alan_cat, fit) |>
  pivot_wider(names_from = alan_cat, values_from = fit) |>
  mutate(advance_mins = (Dark - Light) * 60)

# Summarize onset advancement
adv_summary <- family_advances |>
  group_by(cloud_cat) |>
  summarize(
    mean_adv = mean(advance_mins, na.rm = TRUE),
    sd_adv = sd(advance_mins, na.rm = TRUE),
    median_adv = median(advance_mins, na.rm = TRUE)
  )

adv_summary
```


"Roughly half `(52.4%)` of the diurnal avian families showed extended daily vocal activity windows, characterized by both earlier morning onset..."
```{r}
#| label: calc-extended-windows
coef_panel_a_on_wide |>
  group_by(family) |>
  mutate(
    extended = case_when(
      (Onset < 0 &
        Cessation > 0) ~ "extended window",
      TRUE ~ "not extended"
    )
  ) |>
  ungroup() |>
  count(extended) |>
  mutate(percentage = (n / sum(n)) * 100) ## 52.4% extended (33 fam), 47.6% not extended (30 fam)
```


"Most nocturnal species `(75%)` showed lower vocalization probability in bright compared to dark landscapes"
```{r}
#| label: calc-lower-dets
coef_panel_b_wide |>
  group_by(common_name) |>
  mutate(
    lower_det_bright = case_when(
      any(Bright < Dark) ~ "lower",
      TRUE ~ "higher"
    )
  ) |>
  ungroup() |>
  count(lower_det_bright) |>
  mutate(percentage = (n / sum(n)) * 100) ## 75%, 25%
```