---
title: 107_openmeteo_summarize_data_diurnal.qmd
toc: true
format:
 html:
   embed-resources: true
date: 2026-03-16 # last-modified
date-format: "yyyy-MM-dd"
abstract: This script summarizes weather data from the openmeteo API
          for each BirdWeather station, using 3hr buffers around sunset
          and sunrise times across the study period
---

# Set up environment

```{r}
# ------------ SET UP ENVIRONMENT ------------
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(tidyverse, here, furrr, suncalc)

# Clear envionrment
rm(list = ls())
gc()

# Set working directory
here::i_am("scripts/L1/107_openmeteo_summarize_data_diurnal.qmd")
```


# Diurnal data

The time window for the diurnal analysis is a bit shorter since the response variable is time of earliest detection from sunrise (morning) and time of latest detection from sunset (evening). 
Brent and I decided that we should summarise cloud cover in a 3-hr time window around sunrise and sunset.

So, for each station_id, lat, lon, and date, you'll use suncalc to retrieve sunrise & sunset times.

If, for example, sunrise is at 7:45AM, the hour of sunrise would be 7; you'd filter the weather data to the hours of 6, 7, and 8, and calculate the average. You can use the lubridate::hour() function to quickly grab the hour for a date-time column. 

As a sanity check, you should be getting a table with ~2.8 million rows (if you have wide format with separate columns for sunrise and sunset cloud cover), or ~6 million rows (if you have long format)

You'll then join to the prepped diurnal voc data, joining on station ID and date


**my notes**

* diurnal has `first_onset` and `median_dawn`
  * `first_onset`  = Morning onset: time of first detection - local sunrise 
  * `median_dawn` = Morning median: time of 50% detection - local sunrise


## Read station data

```{r}
# Read stations data
bw_stations <- read_csv(here("data", "L0", "stations_mar2026.csv")) |>
  # Assign batch numbers used for weather data
  mutate(batch_id = (row_number() - 1) %/% 60)

gc()
```

## Summarize cloud cover data

Calculate mean cloud cover in 3-hr time window around sunset and sunrise.

* Round sunset & sunrise to nearest hour
* Filter weather data to 3-hr window
* Calculate mean cover in the two windows for each station_ID, date, [REMOVED: lat/lon]

If, for example, sunrise is at 7:45AM, the hour of sunrise would be 7; you'd filter the weather data to the hours of 6, 7, and 8, and calculate the average. You can use the lubridate::hour() function to quickly grab the hour for a date-time column. 

### batch function

```{r}
# set up batching function

# Function to process a batch and summarize cloud cover data
ProcessCurrentBatch <- function(batch) {
  gc()

  batch_name <- paste0("batch_", as.character(batch), ".csv")
  output_path <- file.path(summarized_weather_path, batch_name)

  # If batch file already exists, skip it
  if (file.exists(output_path)) {
    message(sprintf("Skipping Batch: %d - File already exists.", batch))
    return(NULL)
  }

  # Read in weather current batch
  cat(sprintf("Starting Batch: %d", batch))

  cloud <- read_csv(here(
    "data",
    "L1",
    "updated_batches",
    batch_name
  ))

  # Grab subset of stations for current batch
  subset <- bw_stations |> filter(batch_id == batch)

  # Get sunlight info
  sunlight_data <- cloud |>
    mutate(
      date = as.Date(time_iso8601),
      rounded_time = lubridate::round_date(time_iso8601, unit = "hour")
    ) |>
    # Only keep dates within study extent
    # One day before and after is kept though for summarizing data
    filter(between(date, as.Date("2024-03-31"), as.Date("2025-05-01"))) |>
    # Add sunrise/sunset info
    # Other columns like the hourly time_iso8601 make multiple groups here,
    # so we just grab the first one with [1], since the date/lat/lon are
    # all the same within each grouping of station_id / date
    group_by(station_id, date) |>
    mutate(
      sun_data = getSunlightTimes(
        date = date[1],
        lat = latitude[1],
        lon = longitude[1],
        keep = c("sunrise", "sunset"),
        tz = "UTC"
      ) |>
        select(sunrise, sunset)
    ) |>
    unpack(sun_data) |>
    ungroup()

  # Create sunrise/sunset schedule for each station/lat/lon and date
  schedule <- sunlight_data |>
    distinct(station_id, date, sunrise, sunset) |>
    mutate(
      sr_round = lubridate::round_date(sunrise, unit = "hour"),
      ss_round = lubridate::round_date(sunset, unit = "hour")
    )

  # Create a 3-hr buffer window around each sunrise/sunset
  window <- schedule |>
    reframe(
      # Create the 3-hour windows
      rounded_time = c(
        # Sunrise
        sr_round - hours(1),
        sr_round,
        sr_round + hours(1),
        # Sunset
        ss_round - hours(1),
        ss_round,
        ss_round + hours(1)
      ),
      type = rep(c("sunrise", "sunset"), each = 3),
      # This 'date' is the "Anchor Date" requested in getSunlightTimes
      anchor_date = date,
      .by = c(station_id, date)
    )

  # Join window to station/weather/sunlight data
  # > Adds the rounded sunrise/sunset hours to the station/weather/sunlight dataframe
  # > Don't join by `date` because `rounded_time` includes date
  joined <- sunlight_data |>
    left_join(
      window,
      by = c("station_id", "rounded_time")
    )

  # Now summarize the sunset/sunrise windows for each station/lat/lon/date
  final_summary <- joined |>
    group_by(station_id, anchor_date) |>
    summarize(
      sunrise = first(sunrise),
      sunset = first(sunset),
      across(
        contains(c("cover", "mm", "cm")),
        ~ mean(.x[type == "sunrise"], na.rm = TRUE),
        .names = "sunrise_{.col}"
      ),
      across(
        contains(c("cover", "mm", "cm")) & !starts_with("sunrise"),
        ~ mean(.x[type == "sunset"], na.rm = TRUE),
        .names = "sunset_{.col}"
      ),
      .groups = "drop"
    ) |>
    # Clean up column names with % in them
    rename_with(~ gsub("%", "percent", .x))

  # Write batch to CSV
  write_csv(final_summary, output_path)

  cat(sprintf("Finished Batch: %d", batch))
  rm(cloud, sunlight_data, schedule, window, joined, final_summary)
  gc()
}
```

### run batches

```{r}
# For weather data, each batch has 60 stations

# Create a results directory if it doesn't exist
summarized_weather_path <- here(
  "data",
  "L1",
  "107_openmeteo_summarized_batches_diurnal"
)
if (!file.exists(summarized_weather_path)) {
  dir.create(summarized_weather_path)
}


weather_files <- list.files(
  here("data", "L1", "updated_batches"),
  pattern = "*.csv",
  full.names = TRUE
)

max_batch <- length(weather_files) - 1

# Set up parallel processing
plan(multisession, workers = 16)

# Run the map across all batch IDs
future_walk(
  0:max_batch,
  ~ ProcessCurrentBatch(.x),
  .options = furrr_options(seed = TRUE),
  .progress = TRUE
)
```