---
title: 108_openmeteo_summarize_data_nocturnal.qmd
toc: true
format:
 html:
   embed-resources: true
date: 2026-03-16 # last-modified
date-format: "yyyy-MM-dd"
abstract: This script summarizes weather data from the openmeteo API
          for each BirdWeather station, using 3hr buffers around sunset
          and sunrise times across the study period
---

# Set up environment

```{r}
# ------------ SET UP ENVIRONMENT ------------
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(tidyverse, here, furrr, suncalc, data.table)

# Clear envionrment
rm(list = ls())
gc()

# Set working directory
here::i_am("scripts/L1/108_openmeteo_summarize_data_nocturnal.qmd")
```


# Nocturnal data

We'll want to summarize the cloud cover data per night before joining such that we have a table with one row per station ID and night

The tricky bit is that one night spans two calendar dates. So, you'll want to create a column in the weather data called "night_id". If, for example, we have data starting on March 1st, night 1 would begin at sunset on March 1st and end at sunrise on March 2nd, night 2 would begin at sunset on March 2nd and end at sunrise on March 3rd, etc. L23-50 in the 102_data_prep_calculate_nocturnal_vocal_activity.R script (on the GitHub repo) gives an example of how to do this. Brent  - can you chime in if you made significant modifications to this night_id creation in your noct script?

Once you have that night_id column created, you'll want to group by station id and night_id, and summarise the mean (and maybe standard deviation) as well. As a sanity check, this table should be ~2.8 million rows
You'll then join to the prepped nocturnal voc data, joining on station ID and night ID.

on the nocturnal activity files, you will see night_start (sunset heading into night) and night_end (sunrise following morning). You can use the night_start and night_end to summarise the weather data - that should make things simpler/easier


## Read night_id

Using distinct() on the nocturnal vocs dataset to get a smaller dataset with
just unique stations and nights. We need to later join the summarized weather
data back to the nocturnal vocs.

```{r}
# Get all nocturnal voc files in the folder
nocturnal_files <- list.files(
  here("data", "L0", "activity_measures", "nocturnal"),
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8) # parallel::detectCores() - 1)

# Read all CSVs
vocs_nocturnal <- future_map_dfr(
  nocturnal_files,
  function(file) {
    gc() # clear RAM
    read_csv(file)
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
)

# simplify to save memory
nocturnal <- vocs_nocturnal |>
  distinct(
    station_id,
    night_id,
    night_start,
    night_end,
  )

save(nocturnal, file = here("data", "L1", "108a_night_ids.RData"))

rm(vocs_nocturnal)
```

```{r}
# Load in the unique night_ids
load(here("data", "L1", "108a_night_ids.RData"))
```

## Batching function

```{r}
# set up batching function

# Function to process a batch and summarize cloud cover data
ProcessCurrentBatch <- function(batch) {
  gc()

  batch_name <- paste0("batch_", as.character(batch), ".csv")
  output_path <- file.path(summarized_weather_path, batch_name)

  # If batch file already exists, skip it
  if (file.exists(output_path)) {
    message(sprintf("Skipping Batch: %d - File already exists.", batch))
    return(NULL)
  }

  # Read in weather current batch
  cat(sprintf("Starting Batch: %d", batch))

  cloud <- read_csv(here(
    "data",
    "L1",
    "updated_batches",
    batch_name
  ))

  ids_in_batch <- unique(cloud$station_id)

  # Grab subset of stations in the current batch
  subset <- nocturnal |> filter(station_id %in% ids_in_batch)

  # Join nocturnal data to cloud data?
  joined <- subset |>
    mutate(
      rnd_night_start = lubridate::round_date(night_start, unit = "hour"),
      rnd_night_end = lubridate::round_date(night_end, unit = "hour")
    ) |>
    left_join(
      cloud,
      by = join_by(
        station_id == station_id,
        # Weather time must be greater than or equal to start
        rnd_night_start <= time_iso8601,
        # Weather time must be less than or equal to end
        rnd_night_end >= time_iso8601
      )
    )

  # now summarize weather data
  final_summary <- joined |>
    group_by(station_id, night_id, night_start, night_end) |>
    summarize(
      across(
        contains(c("cover", "_mm", "_cm")),
        ~ mean(.x, na.rm = TRUE),
        .names = "nightly_{.col}"
      ),
      .groups = "drop"
    ) |>
    # Clean up column names with % in them
    rename_with(~ gsub("%", "percent", .x))

  # Write batch to CSV
  write_csv(final_summary, output_path)

  cat(sprintf("Finished Batch: %d", batch))
  rm(cloud, ids_in_batch, subset, joined, final_summary)
  gc()
}
```



## Run batches

```{r}
# For weather data, each batch has 60 stations

# Create a results directory if it doesn't exist
summarized_weather_path <- here(
  "data",
  "L1",
  "108b_openmeteo_summarized_batches_nocturnal"
)
if (!file.exists(summarized_weather_path)) {
  dir.create(summarized_weather_path)
}

weather_files <- list.files(
  here("data", "L1", "updated_batches"),
  pattern = "*.csv",
  full.names = TRUE
)

max_batch <- length(weather_files) - 1

# Set up parallel processing
plan(multisession, workers = 16)

# Run the map across all batch IDs
future_walk(
  0:max_batch,
  ~ ProcessCurrentBatch(.x),
  .options = furrr_options(seed = TRUE),
  .progress = TRUE
)
```

## tinkering - check data

```{r}
# --- Check using data from batch_0.csv

#final_summary <- read_csv(here("data", "summarized_weather_batches_nocturnal", "batch_0.csv"))
#cloud <- read_csv(here("data", "updated_batches", "batch_0.csv"))

# From batch_0.csv, let's check station 720 - looks good!
final_summary |>
  filter(
    station_id == 720,
    between(night_start, as.Date("2024-04-01"), as.Date("2024-04-02"))
  ) |>
  print(width = 500)
# start 04-01 16:06
# end   04-02 04:17
cloud |>
  filter(
    station_id == 720,
    between(time_iso8601, as.Date("2024-04-01"), as.Date("2024-04-03"))
  ) |>
  slice(17:29) |> # matching rnd_night_start and rnd_night_end
  summarise(across(everything(), ~ mean(.)))


# From batch_0.csv, let's check station 972 - looks good!
final_summary |>
  filter(
    station_id == 972,
    between(night_start, as.Date("2024-11-05"), as.Date("2024-11-06"))
  )
# start at 2024-11-05 at 10:50
# end at   2024-11-05 at 23:14
cloud |>
  filter(
    station_id == 972,
    between(time_iso8601, as.Date("2024-11-05"), as.Date("2024-11-07"))
  ) |> #print(n=48)
  slice(12:24) |>
  summarise(across(everything(), ~ mean(.)))


# Check using data from batch_81.csv
final_summary <- read_csv(here(
  "data",
  "L1",
  "108b_openmeteo_summarized_batches_nocturnal",
  "batch_81.csv"
))
cloud <- read_csv(here("data", "L1", "updated_batches", "batch_81.csv"))

# From batch_81.csv, let's check station 1131 - looks good!
final_summary |>
  filter(
    station_id == 1131,
    between(night_start, as.Date("2025-03-02"), as.Date("2025-03-03"))
  ) |>
  print(width = 500)
# start 03-02 00:31
# end   03-02 12:57
cloud |>
  filter(
    station_id == 1131,
    between(time_iso8601, as.Date("2025-03-02"), as.Date("2025-03-03"))
  ) |>
  slice(2:14) |> # matching rnd_night_start and rnd_night_end
  summarise(across(everything(), ~ mean(.))) |>
  print(width = 500)

# From batch_81.csv, let's check station 7909 - looks good!
final_summary |>
  filter(
    station_id == 7909,
    between(night_start, as.Date("2024-10-08"), as.Date("2024-10-09"))
  ) |>
  print(width = 500)
# start 10-08 23:05
# end   10-09 11:42
cloud |>
  filter(
    station_id == 7909,
    between(time_iso8601, as.Date("2024-10-08"), as.Date("2024-10-10"))
  ) |> #print(n=48)
  slice(24:37) |> # matching rnd_night_start and rnd_night_end
  summarise(across(everything(), ~ mean(.))) |>
  print(width = 500)
```