

# Set up environment

```{r}
# ------------ SET UP ENVIRONMENT ------------
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(tidyverse, here, furrr, suncalc, data.table)

# Clear envionrment
rm(list = ls())
gc()

# Set working directory
here::i_am("scripts/L1/109_moon_summarize_intensity.qmd")
```


# Nocturnal data 

Script logic:

Hourly moonlight data was extracted for each station_id.

We had discussed some ideas for summarizing moonlight intensity.
I summarized multiple metrics and we can justify one or pick one with AIC

1. Join moonlight data to night_id (join_by rounded night_start / night_end)
2. Summarize moonlight data for each station_id/night_id. Multiple summaries
were calculated:
  * mean moonlight between night_start and night_end
  * median moonlight between night_start and night_end
  * max moonlight between night_start and night_end
  * min moonlight between night_start and night_end
  * moonlight at approximately "solar midnight"
    * rather than recalculating the exact "solar midnight", I just pull the moonlight value halfway between night_start & night_end. If there an even number of hours, I just take the larger moonlight value.

## Read night_id

```{r}
load(here("data", "L1", "108a_night_ids.RData"))
```


## Batching function

```{r}
# set up batching function

# Function to process a batch and summarize moonlight data
ProcessCurrentBatch <- function(batch) {
  gc()

  batch_name <- paste0("batch_", as.character(batch), ".csv")
  output_path <- file.path(summarized_moonlight_path, batch_name)

  # If batch file already exists, skip it
  if (file.exists(output_path)) {
    message(sprintf("Skipping Batch: %d - File already exists.", batch))
    return(NULL)
  }

  # Read in moonlight current batch
  cat(sprintf("Starting Batch: %d", batch))

  moonlight <- read_csv(here(
    "data",
    "L1",
    "updated_batches_buffer", ################################################ BUFFER
    batch_name
  ))

  ids_in_batch <- unique(moonlight$station_id)

  # Grab subset of stations in the current batch
  subset <- nocturnal |> filter(station_id %in% ids_in_batch)

  # Join nocturnal data to moonlight data?
  joined <- subset |>
    mutate(
      rnd_night_start = lubridate::round_date(night_start, unit = "hour"),
      rnd_night_end = lubridate::round_date(night_end, unit = "hour")
    ) |>
    left_join(
      moonlight,
      by = join_by(
        station_id == station_id,
        # Moonlight time must be greater than or equal to start
        rnd_night_start <= datetime_utc,
        # Moonlight time must be less than or equal to end
        rnd_night_end >= datetime_utc
      )
    )

  # now summarize moonlight data
  final_summary <- joined |>
    group_by(station_id, night_id, night_start, night_end) |>
    summarize(
      # standard summary stats
      mean_moonlight = mean(moonlight, na.rm = TRUE),
      median_moonlight = median(moonlight, na.rm = TRUE),
      max_moonlight = max(moonlight, na.rm = TRUE),
      min_moonlight = min(moonlight, na.rm = TRUE),

      # approx solar midnight
      # Note that there might not be data for the hours of night start & end
      #  because it was not quite yet night, according to moonlight calculation
      #  logic from MoonShineR
      midnight_moonlight = {
        # Find the time halfway between night start and end
        mid = night_start + (night_end - night_start) / 2

        # Calculate time difference for all available moonlight data rows
        diffs = abs(difftime(datetime_utc, mid, units = "hours"))

        # Find lowest time difference (closest to midpoint)
        # This addresses cases where midpoint isn't available because midpoint
        # occurred at a time that still had daylight, according to moonlight calculation logic from MoonShineR
        vals = moonlight[diffs == min(diffs, na.rm = TRUE)]

        # CATCH NA VALUES FROM WRONG NIGHT_IDs & >24-hr NIGHTS ISSUE - resolved :]
        if (length(vals) > 0) {
          # If there is a tie (midpoint is exactly between two hours),
          # then take the max() moonlight value.
          max(vals, na.rm = TRUE)
        } else {
          NA
        }
      },
      .groups = "drop"
    )

  # Write batch to CSV
  write_csv(final_summary, output_path)

  cat(sprintf("Finished Batch: %d", batch))
  rm(moonlight, ids_in_batch, subset, joined, final_summary)
  gc()
}
```



## Run batches

```{r}
# For moonlight data, each batch has 60 stations

# Create a results directory if it doesn't exist
summarized_moonlight_path <- here("data", "L1", "109_moon_summarized_batches")
if (!file.exists(summarized_moonlight_path)) {
  dir.create(summarized_moonlight_path)
}

moonlight_files <- list.files(
  here("data", "L1", "updated_batches_buffer"), ################################################ BUFFER
  pattern = "*.csv",
  full.names = TRUE
)

max_batch <- length(moonlight_files) - 1

# Set up parallel processing
plan(multisession, workers = 16)

# Run the map across all batch IDs
future_walk(
  0:max_batch,
  ~ ProcessCurrentBatch(.x),
  .options = furrr_options(seed = TRUE),
  .progress = TRUE
)
```



# tinkering (old code)

## check data

```{r}
# --- Check using data from batch_0.csv

#final_summary <- read_csv(here("data", "L1","109_moon_summarized_batches", "batch_0.csv"))
moonlight <- read_csv(here(
  "data",
  "L1",
  "updated_batches_buffer",
  "batch_0.csv"
))

# From batch_0.csv, let's check station 720 - looks good!
final_summary |>
  filter(
    station_id == 720,
    between(night_start, as.Date("2024-04-01"), as.Date("2024-04-02"))
  )
# start 04-01 16:06
# end   04-02 04:17
moonlight |>
  filter(
    station_id == 720,
    between(
      datetime_utc,
      as.POSIXct("2024-04-01 16:00:00", tz = "UTC"),
      as.POSIXct("2024-04-02 04:00:00", tz = "UTC")
    )
  ) |>
  summarize(
    mean_moonlight = mean(moonlight, na.rm = TRUE),
    median_moonlight = median(moonlight, na.rm = TRUE),
    max_moonlight = max(moonlight, na.rm = TRUE),
    min_moonlight = min(moonlight, na.rm = TRUE),
    midnight_moonlight = {
      # Find the time halfway between night start and end
      # Note that there might not be data for the hours of night start & end
      # because it was not quite yet night, according to moonlight calculation logic from MoonShineR
      mid = as.POSIXct("2024-04-01 16:00:00", tz = "UTC") +
        (as.POSIXct("2024-04-02 04:00:00", tz = "UTC") -
          as.POSIXct("2024-04-01 16:00:00", tz = "UTC")) /
          2

      # Calculate time difference for all available moonlight data rows
      diffs = abs(difftime(datetime_utc, mid, units = "hours"))

      # Find lowest time difference (closest to midpoint)
      # This addresses cases where midpoint isn't available because midpoint
      # occurred at a time that still had daylight, according to moonlight calculation logic from MoonShineR
      vals = moonlight[diffs == min(diffs, na.rm = TRUE)]

      # CATCH NA VALUES FROM WRONG NIGHT_IDs & >24-hr NIGHTS ISSUE - resolved :]
      if (length(vals) > 0) {
        # If there is a tie (midpoint is exactly between two hours),
        # then take the max() moonlight value.
        max(vals, na.rm = TRUE)
      } else {
        NA
      }
    }
  )


# From batch_0.csv, let's check station 972 - looks good!
final_summary |>
  filter(
    station_id == 972,
    between(night_start, as.Date("2024-11-05"), as.Date("2024-11-06"))
  ) |>
  print(width = 200)
# start at 2024-11-05 at 10:50
# end at   2024-11-05 at 23:14
moonlight |>
  filter(
    station_id == 972,
    between(
      datetime_utc,
      as.POSIXct("2024-11-05 11:00:00", tz = "UTC"),
      as.POSIXct("2024-11-05 23:00:00", tz = "UTC")
    )
  ) |>
  summarize(
    mean_moonlight = mean(moonlight, na.rm = TRUE),
    median_moonlight = median(moonlight, na.rm = TRUE),
    max_moonlight = max(moonlight, na.rm = TRUE),
    min_moonlight = min(moonlight, na.rm = TRUE),
    midnight_moonlight = {
      # Find the time halfway between night start and end
      # Note that there might not be data for the hours of night start & end
      # because it was not quite yet night, according to moonlight calculation logic from MoonShineR
      mid = as.POSIXct("2024-11-05 11:00:00", tz = "UTC") +
        (as.POSIXct("2024-11-05 23:00:00", tz = "UTC") -
          as.POSIXct("2024-11-05 11:00:00", tz = "UTC")) /
          2
      # Calculate time difference for all available moonlight data rows
      diffs = abs(difftime(datetime_utc, mid, units = "hours"))

      # Find lowest time difference (closest to midpoint)
      # This addresses cases where midpoint isn't available because midpoint
      # occurred at a time that still had daylight, according to moonlight calculation logic from MoonShineR
      vals = moonlight[diffs == min(diffs, na.rm = TRUE)]

      # CATCH NA VALUES FROM WRONG NIGHT_IDs & >24-hr NIGHTS ISSUE - resolved :]
      if (length(vals) > 0) {
        # If there is a tie (midpoint is exactly between two hours),
        # then take the max() moonlight value.
        max(vals, na.rm = TRUE)
      } else {
        NA
      }
    }
  )


# Check using data from batch_81.csv
#final_summary <- read_csv(here("data", "L1", "109_moon_summarized_batches","batch_81.csv"))
#moonlight <- read_csv(here("data", "L1", "updated_batches_buffer", "batch_81.csv"))

# From batch_81.csv, let's check station 1131 - looks good!
final_summary |>
  filter(
    station_id == 1131,
    between(night_start, as.Date("2025-03-02"), as.Date("2025-03-03"))
  ) |>
  print(width = 200)
# start 03-02 00:31
# end   03-02 12:57
moonlight |>
  filter(
    station_id == 1131,
    between(
      datetime_utc,
      as.POSIXct("2025-03-02 01:00:00", tz = "UTC"),
      as.POSIXct("2025-03-02 13:00:00", tz = "UTC")
    )
  ) |>
  summarize(
    mean_moonlight = mean(moonlight, na.rm = TRUE),
    median_moonlight = median(moonlight, na.rm = TRUE),
    max_moonlight = max(moonlight, na.rm = TRUE),
    min_moonlight = min(moonlight, na.rm = TRUE),
    midnight_moonlight = {
      # Find the time halfway between night start and end
      # Note that there might not be data for the hours of night start & end
      # because it was not quite yet night, according to moonlight calculation logic from MoonShineR
      mid = as.POSIXct("2025-03-02 01:00:00", tz = "UTC") +
        (as.POSIXct("2025-03-02 13:00:00", tz = "UTC") -
          as.POSIXct("2025-03-02 01:00:00", tz = "UTC")) /
          2
      # Calculate time difference for all available moonlight data rows
      diffs = abs(difftime(datetime_utc, mid, units = "hours"))

      # Find lowest time difference (closest to midpoint)
      # This addresses cases where midpoint isn't available because midpoint
      # occurred at a time that still had daylight, according to moonlight calculation logic from MoonShineR
      vals = moonlight[diffs == min(diffs, na.rm = TRUE)]

      # CATCH NA VALUES FROM WRONG NIGHT_IDs & >24-hr NIGHTS ISSUE - resolved :]
      if (length(vals) > 0) {
        # If there is a tie (midpoint is exactly between two hours),
        # then take the max() moonlight value.
        max(vals, na.rm = TRUE)
      } else {
        NA
      }
    }
  )


# From batch_81.csv, let's check station 7909 - looks good!
final_summary |>
  filter(
    station_id == 7909,
    between(night_start, as.Date("2024-10-08"), as.Date("2024-10-09"))
  ) |>
  print(width = 200)
# start 10-08 23:05
# end   10-09 11:42
moonlight |>
  filter(
    station_id == 7909,
    between(
      datetime_utc,
      as.POSIXct("2024-10-08 23:00:00", tz = "UTC"),
      as.POSIXct("2024-10-09 12:00:00", tz = "UTC")
    )
  ) |>
  summarize(
    mean_moonlight = mean(moonlight, na.rm = TRUE),
    median_moonlight = median(moonlight, na.rm = TRUE),
    max_moonlight = max(moonlight, na.rm = TRUE),
    min_moonlight = min(moonlight, na.rm = TRUE),
    midnight_moonlight = {
      # Find the time halfway between night start and end
      # Note that there might not be data for the hours of night start & end
      # because it was not quite yet night, according to moonlight calculation logic from MoonShineR
      mid = as.POSIXct("2024-10-08 23:00:00", tz = "UTC") +
        (as.POSIXct("2024-10-09 12:00:00", tz = "UTC") -
          as.POSIXct("2024-10-08 23:00:00", tz = "UTC")) /
          2
      # Calculate time difference for all available moonlight data rows
      diffs = abs(difftime(datetime_utc, mid, units = "hours"))

      # Find lowest time difference (closest to midpoint)
      # This addresses cases where midpoint isn't available because midpoint
      # occurred at a time that still had daylight, according to moonlight calculation logic from MoonShineR
      vals = moonlight[diffs == min(diffs, na.rm = TRUE)]

      # --- old logic [start] ---
      # Identify the target hours from midpoint time
      #targets <- c(floor_date(mid, "hour"), ceiling_date(mid, "hour"))

      # Grab the moonlight values for the target hours
      #vals <- moonlight[datetime_utc %in% targets]
      # --- old logic [end] ---

      # CATCH NA VALUES FROM WRONG NIGHT_IDs & >24-hr NIGHTS ISSUE - resolved :]
      if (length(vals) > 0) {
        # If there is a tie (midpoint is exactly between two hours),
        # then take the max() moonlight value.
        max(vals, na.rm = TRUE)
      } else {
        NA
      }
    }
  )
```

## tinker - check ALL data


```{r}
# import from combine_data

all_data_path <- here(
  "data",
  "L2",
  "201d_summarized_weather_data_nocturnal.RData"
)

load(all_data_path)


# --- Check for any NAs
moonlight_nocturnal |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  print(width = 1000)

# - (n=882 rows)
moonlight_nocturnal |> filter(if_any(everything(), is.na)) |> count()

# - Look at dates of NAs and whatnot - NAs dont seem to have any patterns
# (good; not caused by any logic from summarizing and joining weather data to vocs data)
nas <- moonlight_nocturnal |> filter(if_any(everything(), is.na))

# get na stations
ids_in_batch <- unique(nas$station_id)

# Grab subset of na stations
subset <- nocturnal |> filter(station_id %in% ids_in_batch)


summarized_moonlight_path <- here("data", "L1", "109_moon_summarized_batches")

# Combine the list into one big dataframe
summarized_files <- list.files(
  summarized_moonlight_path,
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8)

# Read all CSVs
na_data <- future_map_dfr(
  summarized_files,
  function(file) {
    gc() # clear RAM
    batch_data <- read_csv(file, show_col_types = FALSE)

    stations <- batch_data |>
      filter(station_id %in% subset$station_id) |>
      mutate(file = as.character(file))

    #batch_data[station_id %in% subset$station_id]

    # save if we find some stations
    if (length(stations) > 0) {
      return(stations)
    } else {
      return(NA)
    }
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
) |>
  na.omit()
```


### batch 1 station 5855

```{r}
# From batch_1.csv, let's check station 5855
# identified issue with midpoint logic failing to catch midpoints that [FIXED]
# are NA due to MoonShineR calculations dropping sunlight in moon data [FIXED]
final_summary |>
  filter(
    station_id == 5855,
    between(night_start, as.Date("2024-10-16"), as.Date("2024-10-17"))
  ) |>
  print(width = 200)
# start 2024-10-16 10:34:31
# end   2024-10-16 23:10:34
moonlight |>
  filter(
    station_id == 5855,
    between(
      datetime_utc,
      as.POSIXct("2024-10-16 11:00:00", tz = "UTC"),
      as.POSIXct("2024-10-16 23:00:00", tz = "UTC")
    )
  )
summarize(
  mean_moonlight = mean(moonlight, na.rm = TRUE),
  median_moonlight = median(moonlight, na.rm = TRUE),
  max_moonlight = max(moonlight, na.rm = TRUE),
  min_moonlight = min(moonlight, na.rm = TRUE),
  midnight_moonlight = {
    # Find the time halfway between night start and end
    # Note that there might not be data for the hours of night start & end
    # because it was not quite yet night, according to moonlight calculation logic from MoonShineR
    mid = as.POSIXct("2024-10-16 11:00:00", tz = "UTC") +
      (as.POSIXct("2024-10-16 23:00:00", tz = "UTC") -
        as.POSIXct("2024-10-16 11:00:00", tz = "UTC")) /
        2

    # Calculate time difference for all available moonlight data rows
    diffs = abs(difftime(datetime_utc, mid, units = "hours"))

    # Find lowest time difference (closest to midpoint)
    # This addresses cases where midpoint isn't available because midpoint
    # occurred at a time that still had daylight, according to moonlight calculation logic from MoonShineR
    vals = moonlight[diffs == min(diffs, na.rm = TRUE)]

    # --- old logic [start] ---
    # Identify the target hours from midpoint time
    #targets <- c(floor_date(mid, "hour"), ceiling_date(mid, "hour"))

    # Grab the moonlight values for the target hours
    #vals <- moonlight[datetime_utc %in% targets]
    # --- old logic [end] ---

    # CATCH NA VALUES FROM WRONG NIGHT_IDs & >24-hr NIGHTS ISSUE - resolved :]
    if (length(vals) > 0) {
      # If there is a tie (midpoint is exactly between two hours),
      # then take the max() moonlight value.
      max(vals, na.rm = TRUE)
    } else {
      NA
    }
  }
)
```


### tinker - station 8856

```{r}
# From batch_131.csv, let's check station 8856
uh <- read_csv(here(
  "data",
  "L1",
  "summarized_moonlight_batches",
  "batch_131.csv"
))
# Looks like this station is missing a lot of data...
final_summary |>
  filter(
    station_id == 8856,
    between(night_start, as.Date("2025-04-02"), as.Date("2025-04-03"))
  ) |>
  print(width = 200)
# start 2025-04-02 23:15:47
# end   2025-04-03 11:05:54
# You can see that there is simply moonlight data missing for some reason
moonlight |>
  filter(
    station_id == 8856,
    between(
      datetime_utc,
      as.POSIXct("2025-04-02 23:00:00", tz = "UTC"),
      as.POSIXct("2025-04-03 11:00:00", tz = "UTC")
    )
  )

# If we loko at the lat/lon of this station, it's [90, -119]
# this station is in the middle of the Arctic ocean....
# I also investigated this station in the `get_moonlight_intensity` script.
# Because it's in the middle of the ocean, there was no elevation data for it.
# However NA elevations just default to 0.
# Some dates for "moon_final_lux_nighttime" just show up as NA, which may
# be related to Arctic dates/seasons & winters & auroras and whatnot
# this is why there is data for some dates and not others.
# regardless, we are going to throw out this station.

brent_stations <- read_csv(here(
  "data",
  "L0",
  "stations_mar2026.csv"
)) |>
  filter(station_id == 8856)
```