---
title: 201_combine_data.qmd
subtitle: Clouds modulate light-pollution effects on avian behavior
toc: true
format:
  html:
    embed-resources: false
    cache: false
abstract: This script combines covariate datasets with vocalization activity data
date: 2026-09-12 # last modified
date-format: "yyyy-MM-dd"
---

# Set up environment

```{r}
# ------------ SET UP ENVIRONMENT ------------
if (!require("pacman")) {
  install.packages("pacman")
}

pacman::p_load(tidyverse, here, furrr, suncalc, data.table)

# Clear envionrment
rm(list = ls())
gc()

# Set working directory
here::i_am("scripts/L2/201_combine_data.qmd")
```


# Diurnal data


## Read & Save 

Calculate mean cloud cover in 3-hr time window around sunset and sunrise.

* Round sunset & sunrise to nearest hour
* Filter weather data to 3-hr window
* Calculate mean cover in the two windows for each station_ID, date, lat/lon

```{r}
#| label: summerized-weather-data-diurnal
summarized_weather_path <- here(
  "data",
  "L1",
  "107_openmeteo_summarized_batches_diurnal"
)

# Combine the list into one big dataframe
summarized_files <- list.files(
  summarized_weather_path,
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8)

# Read all CSVs
weather_diurnal <- future_map_dfr(
  summarized_files,
  function(file) {
    gc() # clear RAM
    read_csv(file)
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
)

save(
  weather_diurnal,
  file = here("data", "L2", "summarized_weather_data_diurnal.RData")
)
```


```{r}
#| label: vocs-data-diurnal
# Get all diurnal voc files in the folder
diurnal_files <- list.files(
  here("data", "L0", "activity_measures", "diurnal"),
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8) # parallel::detectCores() - 1)

# Read all CSVs
vocs_diurnal <- future_map_dfr(
  diurnal_files,
  function(file) {
    gc() # clear RAM
    read_csv(file)
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
)

save(
  vocs_diurnal,
  file = here("data", "L2", "activity_measures_diurnal.RData")
)
```


```{r}
#| label: viirs-data
# Read in VIIRS data - pick one to read in
viirs_path <- here("data", "L1", "101_stations_with_VIIRS_2026_03_06.csv")

viirs <- read_csv(viirs_path)

save(
  viirs,
  file = here("data", "L2", "stations_with_VIIRS_2026_03_06.RData")
)
```

## Join data

Join diurnal vocs & weather & VIIRS data

```{r}
#| label: read-data-diurnal
# Quick load data
# Diurnal weather data
load(here("data", "L2", "summarized_weather_data_diurnal.RData"))
# Diurnal vocalizations
load(here("data", "L2", "activity_measures_diurnal.RData"))
# VIIRS data
load(here("data", "L2", "stations_with_VIIRS_2026_03_06.RData"))
```

```{r}
#| label: join-data-diurnal
# Join weather and VIIRS data to diurnal vocs
final_diurnal <- vocs_diurnal |>
  mutate(
    year = year(date),
    month = str_pad(month(date), width = 2, pad = "0")
  ) |>
  left_join(
    viirs,
    by = c(
      # not joining by lat/lon because station_ids are all unique
      # viirs has `lat` and `lon` cols that need to be removed
      "station_id",
      "year",
      "month"
    )
  ) |>
  left_join(
    weather_diurnal,
    by = c(
      "station_id",
      "date" = "anchor_date"
    )
  ) |>
  # ---- Clean up the dataframe for export
  # Change month to an integer instead of a string
  mutate(month = as.integer(month)) |>
  # Reorder columns
  relocate(year, month, .after = year) |>
  # Remove unnecessary columns from VIIRS data
  select(-lat_lon, -lat, -lon)


# Check VIIRS and final was successful
viirs |>
  filter(station_id == 1719 & year == 2024 & month == "04") |>
  print(width = 1000)
final_diurnal |>
  filter(station_id == 1719 & year == 2024 & month == 4) |>
  print(width = 1000)

# Check weather and final join was successfull
weather_diurnal |>
  filter(station_id == 2402 & anchor_date == "2024-04-28") |>
  print(width = 1000)
final_diurnal |>
  filter(station_id == 2402 & date == "2024-04-28") |>
  print(width = 1000)


# --- Check for any NAs
final_diurnal |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  print(width = 1000)

# -  station_id=4979 is missing some weather data
final_diurnal |>
  filter(station_id == 4979) |>
  filter(if_any(everything(), is.na)) |>
  glimpse()

# - Also, station 8856 is missing some weather data
# Due to the issues observed with this station (see nocturnal section for more info)
# station_id=8856 was REMOVED
final_diurnal |>
  filter(station_id == 8856) |>
  glimpse()

# station_id==2321 was missing a few sunrise/set values
final_diurnal |>
  filter(station_id == 2321 & (is.na(sunrise) | is.na(sunset))) |>
  count()

# - REMOVE OBSERVED NA VALUES
final_diurnal <- final_diurnal |>
  filter(
    !(station_id %in% c(4979, 8856)) &
      !(is.na(sunrise) | is.na(sunset))
  )

# Now there are no unsolved cases of NAs :]
final_diurnal |> filter(if_any(everything(), is.na)) |> count()

### REMOVE 7 ADDITIONAL NA VALUES
final_diurnal <- final_diurnal |>
  filter(!(is.na(value)))


#nas <- final_diurnal |> filter(if_any(everything(), is.na))
# sum_nas <- nas |>
#   group_by(station_id, date) |>
#   summarize(across(everything(), ~ sum(is.na(.))))

# A couple stations have some duplicate vocalizations, so we de-dupe them
final_diurnal |>
  filter(category == "ev_ces") |>
  count(station_id, date, common_name, datetime, sort = TRUE) |>
  filter(n > 1)

final_diurnal |> filter(category == "ev_ces") |> nrow()
final_diurnal |>
  filter(category == "ev_ces") |>
  distinct(station_id, date, common_name, datetime, .keep_all = TRUE) |>
  nrow()

# Create different datasets to analyze for modelling
final_diurnal_first_onset <- final_diurnal |>
  distinct(station_id, date, common_name, datetime, .keep_all = TRUE) |>
  filter(category == "first_onset") |>
  pivot_wider(names_from = category, values_from = value) |>
  relocate(first_onset, .before = sunrise)

final_diurnal_ev_ces <- final_diurnal |>
  distinct(station_id, date, common_name, datetime, .keep_all = TRUE) |>
  filter(category == "ev_ces") |>
  pivot_wider(names_from = category, values_from = value) |>
  relocate(ev_ces, .before = sunrise)

###### EXPORT DATASET
save(final_diurnal, file = here("data", "L2", "final_diurnal.RData"))
save(
  final_diurnal_first_onset,
  file = here("data", "L2", "final_diurnal_first_onset.RData")
)
save(
  final_diurnal_ev_ces,
  file = here("data", "L2", "final_diurnal_ev_ces.RData")
)
```

# Nocturnal analysis

```{r}
#| label: clean-env
rm(list = ls())
gc()
```

```{r}
#| label: read-station-data
bw_stations <- read_csv(here("data", "L0", "stations_mar2026.csv")) |>
  # Assign batch numbers used for weather data
  mutate(batch_id = (row_number() - 1) %/% 60)
gc()
```

## Read & Save data


```{r}
#| label: summarized-weather-data-nocturnal
summarized_weather_path <- here(
  "data",
  "L1",
  "108b_openmeteo_summarized_batches_nocturnal"
)
# Combine the list into one big dataframe
summarized_files <- list.files(
  summarized_weather_path,
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8)

# Read all CSVs
weather_nocturnal <- future_map_dfr(
  summarized_files,
  function(file) {
    gc() # clear RAM
    read_csv(file)
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
)

save(
  weather_nocturnal,
  file = here("data", "L2", "summarized_weather_data_nocturnal.RData")
)
```


```{r}
#| label: summarized-moonlight-data
summarized_weather_path <- here("data", "L1", "109_moon_summarized_batches")

# Combine the list into one big dataframe
summarized_files <- list.files(
  summarized_moonlight_path,
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8)

# Read all CSVs
moonlight_nocturnal <- future_map_dfr(
  summarized_files,
  function(file) {
    gc() # clear RAM
    read_csv(file)
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
)

save(
  moonlight_nocturnal,
  file = here("data", "L2", "summarized_moonlight_data_nocturnal.RData")
)
```

```{r}
#| label: vocs-data-nocturnal
# Get all nocturnal voc files in the folder
nocturnal_files <- list.files(
  here("data", "L0", "activity_measures", "nocturnal"),
  pattern = "*.csv",
  full.names = TRUE
)

# Set up parallel processing to read data faster
# Make sure you have enough RAM and cores (adjust worker number as needed)
plan(multisession, workers = 8) # parallel::detectCores() - 1)

# Read all CSVs
vocs_nocturnal <- future_map_dfr(
  nocturnal_files,
  function(file) {
    gc() # clear RAM
    read_csv(file)
  },
  .options = furrr_options(seed = TRUE, scheduling = 1)
)

save(
  vocs_nocturnal,
  file = here("data", "L2", "activity_measures_nocturnal.RData")
)
```

### VIIRS data

This was already done, just load RData object

## Join data

Join nocturnal vocs & weather & moonlight & VIIRS data

```{r}
#| label: read-data-nocturnal
# Nocturnal weather data
load(here("data", "L2", "summarized_weather_data_nocturnal.RData"))
# Moonlight data
load(here("data", "L2", "summarized_moonlight_data_nocturnal.RData"))
# Nocturnal vocalizations
load(here("data", "L2", "activity_measures_nocturnal.RData"))
# VIIRS data
load(here("data", "L2", "stations_with_VIIRS_2026_03_06.RData"))
```


```{r}
#| label: join-data-nocturnal
# reuse some of the diurnal logic here

# Join weather, VIIRS, and moonlight data to nocturnal vocs
final_nocturnal <- vocs_nocturnal |>
  # byebye to station_id=8856 with lat/lon coords in Arctic ocean: [90,-119]
  # We only lose 839 vocs
  filter(station_id != 8856) |>
  mutate(
    #### Use month/year of night_start for joining with VIIRS
    year = year(night_start),
    month = str_pad(month(night_start), width = 2, pad = "0")
  ) |>
  left_join(
    viirs,
    by = c(
      # not joining by lat/lon because station_ids are all unique
      # viirs has `lat` and `lon` cols that need to be removed
      "station_id",
      "year",
      "month"
    )
  ) |>
  left_join(
    weather_nocturnal,
    by = c(
      # These cols were used to get distinct cols for summarizing the data
      # in the weather summary data script
      "station_id",
      "night_id",
      "night_start",
      "night_end"
      #"station_id",
      #"date" = "anchor_date"
    )
  ) |>
  left_join(
    moonlight_nocturnal,
    by = c(
      # These cols were used to get distinct cols for summarizing the data
      # in the weather summary data script
      "station_id",
      "night_id",
      "night_start",
      "night_end"
    )
  ) |>
  # ---- Clean up the dataframe for export
  # Change month to an integer instead of a string
  mutate(month = as.integer(month)) |>
  # Reorder columns
  relocate(year, month, .after = night_end) |>
  # Remove unnecessary columns from VIIRS data
  select(-lat_lon, -lat, -lon)

# Check VIIRS and final was successful
viirs |>
  filter(station_id == 1719 & year == 2024 & month == "04") |>
  print(width = 200)
final_nocturnal |>
  filter(station_id == 1719 & year == 2024 & month == "04") |>
  print(width = 200)

# Check weather and final join was successfull
weather_nocturnal |>
  filter(station_id == 2402 & night_id == "1") |>
  print(width = 1000)
final_nocturnal |>
  filter(station_id == 2402 & night_id == "1") |>
  print(width = 200)

# Check moonlight and final join was successful
moonlight_nocturnal |>
  filter(
    station_id == 8234 & night_id == "23" & night_start >= as.Date("2025-04-24")
  ) |>
  print(width = 1000)
final_nocturnal |>
  filter(
    station_id == 8234 & night_id == "23" & night_start >= as.Date("2025-04-24")
  ) |>
  print(width = 1000)

# --- Check for any NAs
# station_id=8856 was already removed during the data join above
# YIPPEE NO NAs!!
final_nocturnal |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  print(width = 1000)

# - look at NAs, if any
#final_nocturnal |> filter(if_any(everything(), is.na)) |> count()
#nas <- final_nocturnal |> filter(if_any(everything(), is.na))
# sum_nas <- nas |>
#   group_by(station_id, date) |>
#   summarize(across(everything(), ~ sum(is.na(.))))

###### EXPORT DATASET - multiple types for safety
write_rds(final_nocturnal, here("data", "L2", "final_nocturnal.rds"))
save(final_nocturnal, file = here("data", "L2", "final_nocturnal.RData"))
data.table::fwrite(final_nocturnal, here("data", "L2", "final_nocturnal.csv"))
```



# old code

### weather NAs

```{r}
# got some weather NAs - what's up?

# look at diurnal
diurnalnas <- final_diurnal |> filter(if_any(everything(), is.na))

distinctdiurnalnas <- diurnalnas |>
  distinct(station_id, date)

huh <- bw_stations |>
  filter(station_id %in% distinctdiurnalnas$station_id)


#look at nocturnal
distinct_nas <- nas |>
  distinct(
    station_id,
    night_id,
    night_start,
    night_end,
    #.keep_all = TRUE
  )

hmm <- bw_stations |>
  filter(station_id %in% distinct_nas$station_id)
bw_stations
```


```{r}
# Grab unique night_id and night_start and night_end
night_id <- nocturnal |>
  distinct(
    station_id,
    night_id,
    night_start,
    night_end,
    #.keep_all = TRUE
  )
```